О множествах точек на плоскости с целочисленными расстояниями

### 2019-06-02 pps-linear-bound/pps-linear-lower-bound.tex
Language problems
Putting all the conjectures together

\begin{conjecture}
	\label{con:no_optimal_with_edge_1}
	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
	is not optimal.
\end{conjecture}

\begin{theorem}
	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
	and the point $(0,\sqrt{A_n})$,
	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
\end{theorem}
Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar} below,
the points with distance 1 should be $(\pm 1/2, 0)$.

@@

\begin{remark}
	The estimate of Theorem~\ref{thm:main_estimate} is not tight.
	The derivation of tight bounds for the minimum diameter $d(2, n)$
	is still a challenging task for the future~\cite[Section 7]{kurz2008minimum}.
\end{remark}
TODO: should I give the known upper bound here?

@@

>>>>

@@

\begin{remark}
	The estimate of Theorem~\ref{thm:main_estimate} is not tight.
	The derivation of tight bounds for the minimum diameter $d(2, n)$
	is still a challenging task for the future~\cite[Section 7]{kurz2008minimum}.
\end{remark}
The known upper bound is $d(2,n)\leq 2^{c \log n \log \log n}$~\cite{harborth1993upper}.
However, we hope that the present article can give a framework for future better estimates.

\begin{conjecture}
	The approach of Theorem~\ref{thm:main_estimate} can be generalized to higher dimensions.
\end{conjecture}
For the overview of current lower bounds for higher dimensions, we refer the reader to~\cite{nozaki2013lower}.

There are also the following ways to improve the current bound,
based on Lemma~\ref{lem:square_container}.

\begin{problem}
	\label{prb:square_side_length_and_diameter_power}
	Is there a number $\delta < 1$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
	belongs to a square of side length $d^\delta$?
\end{problem}
If the answer for Problem~\ref{prb:square_side_length_and_diameter_power} is affirmative,
then we can build an estimate better than linear.

\begin{problem}
	\label{prb:square_side_length_and_diameter_linear}
	What are the maximal numbers $\delta \leq 1$ and $h \leq 0$
	such that any planar integral point set $M$ with $\operatorname{diam} M = d$
	belongs to a square of side length $\delta d - h$?
\end{problem}

If the answer for Problem~\ref{prb:square_side_length_and_diameter_linear} is not $\delta = 1$ and $h = 0$,
then the linear estimate can be slightly improved.

However, Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}
do not seem to show the way for improving the estimate
due to the following well-known theorem (TODO: refs!).
\begin{theorem}
	\label{thm:rational_set_on_circle}
	There exists a dense subset $P$ of the unit circle such that for any $P_1, P_2 \in P$
	the distance $|P_1 P_2|$ is rational.
\end{theorem}
We conjecture that thorough investigation of existing examples of circular integral point sets
(TODO: refs!)
will lead to the negative answers for the
Problems~\ref{prb:square_side_length_and_diameter_power} and~\ref{prb:square_side_length_and_diameter_linear}.

Another possibility to increase the constant in out lower bound is introduced by the following problem.

\begin{problem}
	\label{prb:minimal_shape}
	What is the minimal shape $\mathcal{S}$ such that any planar integral point set $M$ with $\operatorname{diam} M = d$
	belongs to $\mathcal{S}$?
\end{problem}
This problem in another generalization of Lemma~\ref{lem:square_container}.
We know that $\mathcal{S}$ is not a circle of diameter $d$
(the counterexample is a right triangle of side length $d$)
and we conjecture that $\mathcal{S}$ is not a Reuleaux triangle of width $d$
(the probable counterexample belongs to a circle of diameter $d$, see Theorem~\ref{thm:rational_set_on_circle}).

The solution of Problem~\ref{prb:minimal_shape} (if ever found) will lead to sophisticated packing problem,
something like ``How to put $n$ points in a convex hull of concentric circle and Reuleaux triangle?''
Obviously, Problem~\ref{prb:minimal_shape} can be generalized to higher dimensions.

For $n=3,4,5$, there are optimal sets in $\mathfrak{M}(2,n)$ that contain distance 1
(see Remarks~\ref{rem:planar_set_with_minimeter_1_of_3_points} and~\ref{rem:planar_set_with_minimeter_1_of_4_and_5_points}).
However, we can put forward the following conjecture.
\begin{conjecture}
	\label{con:no_optimal_with_edge_1}
	Every set $M\in\mathfrak{M}(2,n)$, $n\geq 6$, such that for some $M_1,M_2$ equality $|M_1 M_2|=1$ holds,
	is not optimal.
\end{conjecture}

\begin{theorem}
	For every $12 \leq n \leq 122$, there exist a facher optimal sets $M_n\in\mathfrak{M}(2,n)$
	that consist of $n-1$ points $\{ (b_1,0), ..., (b_{n-1},0)\}$
	and the point $(0,\sqrt{A_n})$,
	where $b_1,...,b_{n-1}, A_n$ are integer and $\sqrt{A_n}\notin\mathbb{N}$.
\end{theorem}
Such a set can not contain distance 1: due to Theorem~\ref{thm:minimeter_1_planar},
the points with distance 1 should be $(\pm 1/2, 0)$.

When we talk about integral point sets containing distance 1, another problem arises.

\begin{problem}
	For $M\in\mathfrak{M}(m,n)$, how many times can the distance 1 occur in $M$?
\end{problem}
The answer for $m=2$ is given by Theorem~\ref{thm:minimeter_1_planar}.
For greater $m$, we will be brave enough to put forward
\begin{conjecture}
	Consider $M\in\mathfrak{M}(m,n)$, $M_1,M_2\in M$, $|M_1 M_2| =1$.
	Then either $n = m+1$ or $M$ is a ``blowup'' of a facher set and distance 1 occurs in $m$
	not more than $1+\frac{m^2 - m}{2}$ times.
\end{conjecture}
Such as estimate is based on the quantity of edges in $(m-1)$-dimensional simplex.

@@

	\item
		Observation vs. Proposition

	\item
		estimate vs. bound
